---
title: "Prediction Assignment Writeup"
author: ""
date: "Januar 2016"
output: html_document
---

```{r}
set.seed(323)
library(caret)
```

The training and test data were downloaded and loaded in to R. 
```{r}
#download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",
#                     destfile = "pml-training.csv")
#download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
#                           destfile = "pml-testing.csv")
pmlTrain <- read.csv(file="pml-training.csv")
pmlTest <- read.csv(file="pml-testing.csv")
```

The data contains a number of values that appear to be Excel error strings. These values were replaced with an NA value.
```{r}
pmlTrain[pmlTrain=="#DIV/0!"] <- NA
```

Many variables in the data contain almost no actual values. These variables were removed.
```{r}
pmlTrain <- pmlTrain[,colSums(is.na(pmlTrain))==0]
```

Several other variables were removed that would not contribute to the analysis. 
```{r}
pmlTrain <- pmlTrain[,-c(1:7)]
```

The remaining variables are listed below.
```{r}
names(pmlTrain)
```

The data is partitioned into a training and validation set. 
```{r}
inTrain <- createDataPartition(y = pmlTrain$classe, p = 3/4)[[1]]
training <- pmlTrain[inTrain,]
validation <- pmlTrain[-inTrain,]
```

No remaining variabls appear to have near zero variance.
```{r}
nearZeroVar(training)
```

The first method implemented was a single tree, but the accuracy of the model is rather low.
```{r}
rpmodel <- train(classe~., data=training, method="rpart")
rpmodel$results
```

The second method implemented was random forests, and the accuracy was much higher than the first model.
```{r}
rfmodel <- train(classe~., data = training, method = "rf", 
                 trControl = trainControl( method="cv",number=4,allowParallel = TRUE))
rfmodel$results
```

The random forests model prediction on the validation set was also rather accurate. The out-of-sample error from the cross validation set is 0.0071.
```{r}
rfpre <- predict(rfmodel, validation)
confusionMatrix(rfpre,validation$classe)
```

The random forests model was applied to the test set and the predictions are provided below. 
```{r}
predict(rfmodel, pmlTest)
```


